---
title: "plotting"
output: html_document
date: "2024-10-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(grid)
#library(gridExtra)
library(ggpubr)
library(cowplot)
library(tidyverse)
library(data.table)
library(patchwork)
library(ggthemes)
library(extrafont)
library(remotes)
library(ggtext)
#remotes::install_version("Rttf2pt1", version = "1.3.8")
#extrafont::font_import()
```

```{r}
mx_data <- read_csv("~/flu_imprinting/data/mx_data_final.csv", show_col_types = FALSE)
```

want continuous lines without needing to interpolate
```{r}
mx_data <- mx_data %>%
  group_by(year_recode) %>%
  # min and max birth_years with non-NA points for each year
  mutate(
    min_birth_year = min(birth_year[!is.na(log_flu_mx)], na.rm = TRUE),
    max_birth_year = max(birth_year[!is.na(log_flu_mx)], na.rm = TRUE),
    # Set estimates and CIs to NA outside the range of points for each year
    log_flu_mx_est = if_else(birth_year < min_birth_year | birth_year > max_birth_year, NA_real_, log_flu_mx_est),
    log_flu_mx_upper_CI = if_else(birth_year < min_birth_year | birth_year > max_birth_year, NA_real_, log_flu_mx_upper_CI),
    log_flu_mx_lower_CI = if_else(birth_year < min_birth_year | birth_year > max_birth_year, NA_real_, log_flu_mx_lower_CI)
  ) %>%
  ungroup() %>%
  dplyr::select(-min_birth_year, -max_birth_year)

mx_data <- mx_data %>% 
  mutate(imprinted_strain = 
           case_when(
             imprinted_strain == "naive" | imprinted_strain == "ambiguous" ~ "ambiguous/naive",
             TRUE ~ imprinted_strain))
```

Change rates to "per 100,000" for better interpretability
```{r}
mx_data$log_flu_mx_est_p100000 <- log(exp(mx_data$log_flu_mx_est) * 100000)
mx_data$log_flu_mx_p100000 <- log(mx_data$flu_mx * 100000)
mx_data$upper_est_p100000 <- log(exp(mx_data$log_flu_mx_upper_CI) * 100000)
mx_data$lower_est_p100000 <- log(exp(mx_data$log_flu_mx_lower_CI) * 100000)

mx_data$log_flu_mx_p100000[is.infinite(mx_data$log_flu_mx_p100000)] <- NA
```

factor to order legend
```{r}
mx_data$imprinted_strain <- factor(mx_data$imprinted_strain, 
                         levels = c("pH1N1", "H1N1_alpha", "H1N1_beta", 
                                    "H1N1_gamma", "H2N2", "H3N2", "H1N1", "H1N1pdm09", "ambiguous/naive"))
```

various seasons pre 2009
```{r}
p1 = ggplot() +
   geom_ribbon(data = mx_data[mx_data$year_recode == 1974, ], 
              aes(x = birth_year, ymin = lower_est_p100000, ymax = upper_est_p100000), 
              fill = "grey80", alpha = 0.7) +
  geom_line(data = mx_data[mx_data$year_recode == 1974, ], aes(x = birth_year, y = log_flu_mx_est_p100000), na.rm = TRUE) +
  geom_point(data = mx_data[mx_data$year_recode == 1974, ],
             aes(x = birth_year, y = log_flu_mx_p100000, color = imprinted_strain), alpha = 0.6, size = 2, na.rm = TRUE) +
  scale_color_manual(name = "Imprinted strain", 
                          values = c(pH1N1 = "#ff6698",
                                H1N1_alpha = "#b70000", 
                                H1N1_beta = "#ff8900", 
                                H1N1_gamma = "#ECC905", 
                                H2N2 = "#95C90F", 
                                H3N2 = "#3683C3",
                                "ambiguous/naive" = "#989898"),
                     labels = c("ambiguous/naive" = "Ambiguous/Naïve",
                                pH1N1 = bquote(pH1N1["\u03b1"]),
                                H1N1_alpha = bquote(H1N1["\u03b1"]),
                                H1N1_beta = bquote(H1N1["\u03b2"]),
                                H1N1_gamma = bquote(H1N1["\u03b3"]))) +
  labs(x = "", y = "Log Influenza mortality rate") +
  scale_x_reverse(
    breaks = seq(1970, 1850, by = -20),
    sec.axis = sec_axis(
      transform = ~.,
      breaks = seq(1974, 1974-108, by = -20),
      labels = c(0, 20, 40, 60, 80, 100),
      name = "Age")) +
  theme_bw() +
  theme(text = element_text(size=18, family = "Helvetica"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.margin=unit(c(0.1, 0, -0.5, 0.5),"lines"),
        axis.title=element_text(size=14)) +
  scale_y_continuous(limits = c(-5, 6)) +
  geom_label(aes(x = 1958, y = 4.5, label = "1974-1975\nH3N2", family = "Helvetica"), fill = "white", size = 4.5)

p2 = ggplot() +
  geom_ribbon(data = mx_data[mx_data$year_recode == 1981, ], 
              aes(x = birth_year, ymin = lower_est_p100000, ymax = upper_est_p100000), 
              fill = "grey80", alpha = 0.7) +
  geom_line(data = mx_data[mx_data$year_recode == 1981, ], aes(x = birth_year, y = log_flu_mx_est_p100000), na.rm = TRUE) +
  geom_point(data = mx_data[mx_data$year_recode == 1981, ],
             aes(x = birth_year, y = log_flu_mx_p100000, color = imprinted_strain), alpha = 0.6, size = 2, na.rm = TRUE) +
  scale_color_manual(values = c(pH1N1 = "#ff6698", H1N1_alpha = "#b70000", H1N1_beta = "#ff8900", H1N1_gamma = "#ECC905", 
                                H2N2 = "#95C90F", H3N2 = "#3683C3", "ambiguous/naive" = "#989898")) +
      labs(x = "", y = "") +
      scale_x_reverse(
        breaks = seq(1980, 1860, by = -20),
        sec.axis = sec_axis(
          transform = ~.,
          breaks = seq(1981, 1981-108, by = -20),
          labels = c(0, 20, 40, 60, 80, 100),
          name = "Age")) +
  theme_bw() +
  theme(text = element_text(size=18, family = "Helvetica"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.margin=unit(c(0.1, 0, -0.5, 0),"lines"),
        axis.title=element_text(size=14)) + 
  scale_y_continuous(limits = c(-5, 6)) +
  geom_label(aes(x = 1963.5, y = 4.5, label = "1981-1982\nInfluenza B", family = "Helvetica"), fill = "white", size = 4.5)


p3 = ggplot() +
  geom_ribbon(data = mx_data[mx_data$year_recode == 1990, ], 
              aes(x = birth_year, ymin = lower_est_p100000, ymax = upper_est_p100000), 
              fill = "grey80", alpha = 0.7) +
  geom_line(data = mx_data[mx_data$year_recode == 1990, ], aes(x = birth_year, y = log_flu_mx_est_p100000), na.rm = TRUE) +
  geom_point(data = mx_data[mx_data$year_recode == 1990, ],
             aes(x = birth_year, y = log_flu_mx_p100000, color = imprinted_strain), alpha = 0.6, size = 2, na.rm = TRUE) +
  scale_color_manual(values = c(pH1N1 = "#ff6698", H1N1_alpha = "#b70000", H1N1_beta = "#ff8900", H1N1_gamma = "#ECC905", 
                                H2N2 = "#95C90F", H3N2 = "#3683C3", "ambiguous/naive" = "#989898")) +
      labs(x = "Birth year", y = "Log Influenza mortality rate") +
      scale_x_reverse(
        breaks = seq(1990, 1990-108, by = -20),
        sec.axis = sec_axis(
          transform = ~.,
          breaks = seq(1990, 1990-108, by = -20),
          labels = c(0, 20, 40, 60, 80, 100),
          name = "")) +
  theme_bw() +
  theme(text = element_text(size=18, family = "Helvetica"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.margin=unit(c(-0.5, 0, 0.1, 0.5), "lines"),
        axis.title=element_text(size=14)) + 
  scale_y_continuous(limits = c(-5, 6)) +
  geom_label(aes(x = 1962.5, y = 4.5, label = "1990-1991\nNo dominant strain", family = "Helvetica"), fill = "white", size = 4.5)

p4 = ggplot() +
  geom_ribbon(data = mx_data[mx_data$year_recode == 2000, ], 
              aes(x = birth_year, ymin = lower_est_p100000, ymax = upper_est_p100000), 
              fill = "grey80", alpha = 0.7) +
  geom_line(data = mx_data[mx_data$year_recode == 2000, ], aes(x = birth_year, y = log_flu_mx_est_p100000), na.rm = TRUE) +
  geom_point(data = mx_data[mx_data$year_recode == 2000, ],
             aes(x = birth_year, y = log_flu_mx_p100000, color = imprinted_strain), alpha = 0.6, size = 2, na.rm = TRUE) +
  scale_color_manual(values = c(pH1N1 = "#ff6698", H1N1_alpha = "#b70000", H1N1_beta = "#ff8900", H1N1_gamma = "#ECC905", 
                                H2N2 = "#95C90F", H3N2 = "#3683C3", "ambiguous/naive" = "#989898")) +
      labs(x = "Birth year", y = "") +
      scale_x_reverse(
        breaks = seq(2000, 2000-108, by = -20),
        sec.axis = sec_axis(
          transform = ~.,
          breaks = seq(2000, 2000-108, by = -20),
          labels = c(0, 20, 40, 60, 80, 100),
          name = "")) +
  theme_bw() +
  theme(text = element_text(size=18, family = "Helvetica"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.margin=unit(c(-0.5, 0, 0.1, 0), "lines"),
        axis.title=element_text(size=14)) + 
  scale_y_continuous(limits = c(-5, 6)) +
  geom_label(aes(x = 1984.5, y = 4.5, label = "2000-2001\nH1N1", family = "Helvetica"), fill = "white", size = 4.5)
```

```{r}
grid1 = ggarrange(
    p1, p2, p3, p4,
    ncol = 2, nrow = 2,
    common.legend = TRUE, legend = "right",
    widths = c(1, 1, 1),
    heights = c(1, 1, 1),
    font.label = list(size = 18, family = "Helvetica")) +
  theme(plot.margin = unit(c(0, 0.25, 0, 0), "lines"))


ggsave(filename = "~/flu_imprinting/plots/final_figs/various_seasons.png", 
       plot = grid1, height = 6, width = 10, dpi = 800, bg='#ffffff')
```

various seasons starting w/ 2009
```{r}
p5 = ggplot() +
 geom_ribbon(data = mx_data[mx_data$year_recode == 2009, ], 
              aes(x = birth_year, ymin = lower_est_p100000, ymax = upper_est_p100000), 
              fill = "grey80", alpha = 0.7) +
  geom_line(data = mx_data[mx_data$year_recode == 2009, ], aes(x = birth_year, y = log_flu_mx_est_p100000), na.rm = TRUE) +
  geom_point(data = mx_data[mx_data$year_recode == 2009, ],
             aes(x = birth_year, y = log_flu_mx_p100000, color = imprinted_strain), alpha = 0.6, size = 2, na.rm = TRUE) +
  scale_color_manual(name = "Imprinted strain", 
                          values = c(pH1N1 = "#ff6698",
                                H1N1_alpha = "#b70000", 
                                H1N1_beta = "#ff8900", 
                                H1N1_gamma = "#ECC905", 
                                H2N2 = "#95C90F", 
                                H3N2 = "#3683C3",
                                "ambiguous/naive" = "#989898"),
                     labels = c("ambiguous/naive" = "Ambiguous/Naïve",
                                pH1N1 = bquote(pH1N1["\u03b1"]),
                                H1N1_alpha = bquote(H1N1["\u03b1"]),
                                H1N1_beta = bquote(H1N1["\u03b2"]),
                                H1N1_gamma = bquote(H1N1["\u03b3"]))) +
      labs(x = "", y = "Log Influenza mortality rate") +
      scale_x_reverse(
        breaks = seq(2000, 1880, by = -20),
        sec.axis = sec_axis(
          transform = ~.,
          breaks = seq(2009, 2009-108, by = -20),
          labels = c(0, 20, 40, 60, 80, 100),
          name = "Age")) +
  theme_bw() +
  theme(text = element_text(size=18, family = "Helvetica"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.margin=unit(c(0.1, 0, -0.5, 0.5),"lines"),
        axis.title=element_text(size=14)) + 
  scale_y_continuous(limits = c(-4.7, 5.2), n.breaks = 6) +
  geom_label(aes(x = 1991, y = 3.8, label = "2009-2010\nH1N1pdm09", family = "Helvetica"), fill = "white", size = 4.5)

p6 = ggplot() +
geom_ribbon(data = mx_data[mx_data$year_recode == 2012, ], 
              aes(x = birth_year, ymin = lower_est_p100000, ymax = upper_est_p100000), 
              fill = "grey80", alpha = 0.7) +
  geom_line(data = mx_data[mx_data$year_recode == 2012, ], aes(x = birth_year, y = log_flu_mx_est_p100000), na.rm = TRUE) +
  geom_point(data = mx_data[mx_data$year_recode == 2012, ],
             aes(x = birth_year, y = log_flu_mx_p100000, color = imprinted_strain), alpha = 0.6, size = 2, na.rm = TRUE) +
  scale_color_manual(values = c(pH1N1 = "#ff6698", H1N1_alpha = "#b70000", H1N1_beta = "#ff8900", H1N1_gamma = "#ECC905", 
                                H2N2 = "#95C90F", H3N2 = "#3683C3", "ambiguous/naive" = "#989898")) +
      labs(x = "", y = "") +
      scale_x_reverse(
        breaks = seq(2010, 1900, by = -20),
        sec.axis = sec_axis(
          transform = ~.,
          breaks = seq(2012, 2012-108, by = -20),
          labels = c(0, 20, 40, 60, 80, 100),
          name = "Age")) +
  theme_bw() +
  theme(text = element_text(size=18, family = "Helvetica"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.margin=unit(c(0.1, 0, -0.5, 0),"lines"),
        axis.title=element_text(size=14)) + 
  scale_y_continuous(limits = c(-4.7, 5.2), n.breaks = 6) +
  geom_label(aes(x = 1997, y = 3.8, label = "2012-2013\nH3N2", family = "Helvetica"), fill = "white", size = 4.5)

p7 = ggplot() +
geom_ribbon(data = mx_data[mx_data$year_recode == 2013, ], 
              aes(x = birth_year, ymin = lower_est_p100000, ymax = upper_est_p100000), 
              fill = "grey80", alpha = 0.7) +
  geom_line(data = mx_data[mx_data$year_recode == 2013, ], aes(x = birth_year, y = log_flu_mx_est_p100000), na.rm = TRUE) +
  geom_point(data = mx_data[mx_data$year_recode == 2013, ],
             aes(x = birth_year, y = log_flu_mx_p100000, color = imprinted_strain), alpha = 0.6, size = 2, na.rm = TRUE) +
  scale_color_manual(values = c(pH1N1 = "#ff6698", H1N1_alpha = "#b70000", H1N1_beta = "#ff8900", H1N1_gamma = "#ECC905", 
                                H2N2 = "#95C90F", H3N2 = "#3683C3", "ambiguous/naive" = "#989898")) +
      labs(x = "Birth year", y = "Log Influenza mortality rate") +
      scale_x_reverse(
        breaks = seq(2010, 1900, by = -20),
        sec.axis = sec_axis(
          transform = ~.,
          breaks = seq(2013, 2013-108, by = -20),
          labels = c(0, 20, 40, 60, 80, 100),
          name = "")) +
  theme_bw() +
  theme(text = element_text(size = 18, family = "Helvetica"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.margin=unit(c(-0.5, 0, 0.1, 0.5), "lines"),
        axis.title=element_text(size=14)) + 
  scale_y_continuous(limits = c(-4.7, 5.2), n.breaks = 6) +
  geom_label(aes(x = 1995, y = 3.8, label = "2013-2014\nH1N1pdm09", family = "Helvetica"), fill = "white", size = 4.5)

p8 = ggplot() +
geom_ribbon(data = mx_data[mx_data$year_recode == 2020, ], 
              aes(x = birth_year, ymin = lower_est_p100000, ymax = upper_est_p100000), 
              fill = "grey80", alpha = 0.7) +
  geom_line(data = mx_data[mx_data$year_recode == 2020, ], aes(x = birth_year, y = log_flu_mx_est_p100000), na.rm = TRUE) +
  geom_point(data = mx_data[mx_data$year_recode == 2020, ],
             aes(x = birth_year, y = log_flu_mx_p100000, color = imprinted_strain), alpha = 0.6, size = 2, na.rm = TRUE) +
  scale_color_manual(values = c(pH1N1 = "#ff6698", H1N1_alpha = "#b70000", H1N1_beta = "#ff8900", H1N1_gamma = "#ECC905", 
                                H2N2 = "#95C90F", H3N2 = "#3683C3", "ambiguous/naive" = "#989898")) +
      labs(x = "Birth year", y = "") +
      scale_x_reverse(
        breaks = seq(2020, 2020-108, by = -20),
        sec.axis = sec_axis(
          transform = ~.,
          breaks = seq(2020, 2020-108, by = -20),
          labels = c(0, 20, 40, 60, 80, 100),
          name = "")) +
  theme_bw() +
  theme(text = element_text(size = 18, family = "Helvetica"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.margin=unit(c(-0.5, 0, 0.1, 0), "lines"),
        axis.title=element_text(size=14)) + 
  scale_y_continuous(limits = c(-4.7, 5.2), n.breaks = 6) +
  geom_label(aes(x = 1994, y = 3.8, label = "2020-2021\nNo dominant strain", family = "Helvetica"), 
             fill = "white", size = 4.5)
```

```{r}
grid2 = ggarrange(
    p5, p6, p7, p8,
    ncol = 2, nrow = 2,
    common.legend = TRUE, legend = "right",
    widths = c(1, 1, 1),
    heights = c(1, 1, 1)) +
  theme(plot.margin = unit(c(0, 0.25, 0, 0), "lines"))

ggsave(filename = "~/flu_imprinting/plots/final_figs/various_seasons_post_2009.png", 
       plot = grid2, height = 6, width = 10, dpi = 800, bg='#ffffff')
```


grid of all plots for supplementary text
```{r}
arrange_plots <- function(df, obs_col, est_col, plot_years, breaks, ylim, file) {
  df_filtered <- df %>% filter(year_recode %in% plot_years)
  
  plot_list <- lapply(seq_along(plot_years), function(idx) {
    i <- plot_years[idx]
    df_year <- df_filtered %>% filter(year_recode == i)
    
    y_label <- ifelse(idx %% 3 == 1, "Log influenza mortality rate", "")  # only on the left side

    ggplot(df_year) +
      geom_ribbon(aes(x = birth_year, ymin = lower_est_p100000, ymax = upper_est_p100000), fill = "grey80", alpha = 0.7) +
      geom_line(aes(x = birth_year, y = {{est_col}}), na.rm = TRUE) +
      geom_point(aes(x = birth_year, y = {{obs_col}}, color = imprinted_strain), alpha = 0.6, na.rm = TRUE) +
      scale_color_manual(name = "Imprinted strain", 
                         values = c(pH1N1 = "#ff6698", H1N1_alpha = "#b70000", H1N1_beta = "#ff8900", 
                                    H1N1_gamma = "#ECC905", H2N2 = "#95C90F", H3N2 = "#3683C3", 
                                    "ambiguous/naive" = "#989898"),
                         labels = c("ambiguous/naive" = "Ambiguous/Naïve",
                                    pH1N1 = bquote(pH1N1["\u03b1"]),
                                    H1N1_alpha = bquote(H1N1["\u03b1"]),
                                    H1N1_beta = bquote(H1N1["\u03b2"]),
                                    H1N1_gamma = bquote(H1N1["\u03b3"])),
                         guide = guide_legend(ncol = 1)) +
      labs(title = ifelse(i != 2008, 
                          paste0("‎ ", df_year$season[1], ", ", df_year$circulating_strain[1]), 
                          "‎ 2008-2009, H1N1*"),
           x = "Birth year", y = y_label) +
      scale_x_reverse(
        breaks = seq(i - (i %% 5), 1860, by = -20),
        sec.axis = sec_axis(
          transform = ~.,
          breaks = seq(i, i-108, by = -20),
          labels = c(0, 20, 40, 60, 80, 100),
          name = "Age")) +
      scale_y_continuous(breaks = breaks) +
      theme_bw() +
      theme(text = element_text(size=16, family = "Helvetica"),
            plot.title = element_text(vjust = -16, size = 17),
            panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            panel.background = element_blank(),
            axis.title=element_text(size=14),
            legend.key.size = unit(0.7, "cm"),
            legend.text = element_text(size = 15),
            legend.title = element_text(size = 18),
            #legend.box.spacing = unit(0, "cm"), 
            plot.margin = unit(c(-0.75,0.15,0.2,-0.6), "cm")) +
      guides(color = guide_legend(override.aes = list(size=2.5))) +
      coord_cartesian(ylim = ylim)
  })
  
  get_legend <- function(p) {
    g <- ggplotGrob(p)
    legend <- g$grobs[[which(sapply(g$grobs, function(x) x$name) == "guide-box")]]
  }
  
  legend <- get_legend(plot_list[[3]] + theme(legend.position = "right"))
  
  plot_list <- lapply(plot_list, function(p) p + theme(legend.position = "none"))

  grid <- do.call(cowplot::plot_grid, c(plot_list, list(legend, nrow = 3, ncol = 3, align = "hv", axis = "tblr")))
  
  fig <- ggdraw() + draw_plot(grid, x = 0.03, y = 0, width = 0.97, height = 1)

  ggsave(filename = file, plot = fig, height = 10, width = 10, bg='#ffffff', dpi = 800)
}
```

```{r}
arrange_plots(mx_data, log_flu_mx_p100000, log_flu_mx_est_p100000, 1968:1975, c(-2, 0, 2, 4, 6), c(-3.5, 6.7),
              "~/flu_imprinting/plots/final_figs/LC_68-75.png")

arrange_plots(mx_data, log_flu_mx_p100000, log_flu_mx_est_p100000, 1976:1983, c(-4, -2, 0, 2, 4, 6), c(-3.8, 6.6),
              "~/flu_imprinting/plots/final_figs/LC_76-83.png")

arrange_plots(mx_data, log_flu_mx_p100000, log_flu_mx_est_p100000, 1984:1991, c(-4, -2, 0, 2, 4, 6), c(-3.8, 6.5),
              "~/flu_imprinting/plots/final_figs/LC_84-91.png")

arrange_plots(mx_data, log_flu_mx_p100000, log_flu_mx_est_p100000, 1992:1999, c(-4, -2, 0, 2, 4, 6), c(-4, 6.45),
              "~/flu_imprinting/plots/final_figs/LC_92-99.png")

arrange_plots(mx_data, log_flu_mx_p100000, log_flu_mx_est_p100000, 2000:2007, c(-4, -2, 0, 2, 4, 6), c(-4.6, 5.7),
              "~/flu_imprinting/plots/final_figs/LC_00-07.png")

arrange_plots(mx_data, log_flu_mx_p100000, log_flu_mx_est_p100000, 2008:2015, c(-4, -2, 0, 2, 4, 6), c(-4.15, 6.25),
              "~/flu_imprinting/plots/final_figs/LC_08-15.png")

arrange_plots(mx_data, log_flu_mx_p100000, log_flu_mx_est_p100000, 2016:2020, c(-4, -2, 0, 2, 4, 6), c(-4.7, 6.2),
              "~/flu_imprinting/plots/final_figs/LC_16-20.png")
```


flu mortality residuals plot
```{r}
p1 = ggplot(mx_data,
       aes(x = year_recode, y = age_recode, fill = residuals)) +
  geom_tile() +
  scale_fill_gradient2(low = "darkblue", mid = "white", high = "red", midpoint = 0, na.value = "transparent", name = "Excess \nmortality", limits = c(-5.5, 5.5)) +
  labs(x= "Influenza season", y = "Age", tag = "Lower \nmortality") +
  theme_minimal() + 
  theme(plot.tag.position = c(0.8875, 0.375),
        plot.tag = element_text(hjust = 0, size = 15),
        axis.text.x = element_text(angle = 60, hjust = 1, size = 15),
        axis.text.y = element_text(size = 15),
        text = element_text(size = 15, family = "Helvetica"),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=1)) + 
  scale_x_continuous(expand = c(0, 0), breaks = seq(1970, 2020, by = 5), 
                     labels = c("1970", "1975", "1980", "1985", "1990", "1995", "2000",
                                "2005",  "2010", "2015", "2020")) + 
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10))

ggsave(filename = "~/flu_imprinting/plots/final_figs/flu_residuals_plot.png", plot = p1, height = 6, width = 8, bg='#ffffff', dpi = 800)
```

all cause residuals plot
```{r}
p_all <- ggplot(mx_data,
       aes(x = year_recode, y = age_recode, fill = all_cause_residuals)) +
  geom_tile() +
  scale_fill_gradient2(low = "darkblue", mid = "white", high = "red", midpoint = 0, na.value = "transparent", name = "Excess \nmortality", limits = c(-5.6, 5.6)) +
  labs(x= "Flu season", y = "Age", tag = "Lower \nmortality") +
  theme_minimal() + 
   theme(plot.tag.position = c(0.8875, 0.375),
        plot.tag = element_text(hjust = 0, size = 15),
        axis.text.x = element_text(angle = 60, hjust = 1, size = 15),
        axis.text.y = element_text(size = 15),
        text = element_text(size = 15, family = "Helvetica"),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=1)) + 
  scale_x_continuous(expand = c(0, 0), breaks = seq(1970, 2020, by = 5), 
                     labels = c("1970", "1975", "1980", "1985", "1990", "1995", "2000",
                                "2005",  "2010", "2015", "2020")) + 
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10))


ggsave(filename = "~/flu_imprinting/plots/final_figs/all_cause_residuals_plot.png", plot = p_all, height = 6, width = 8, bg='#ffffff')
```

pancreatic cancer residuals plot
```{r}
p_panc <- ggplot(mx_data,
       aes(x = year_recode, y = age_recode, fill = pancreatic_residuals)) +
  geom_tile() +
  scale_fill_gradient2(low = "darkblue", mid = "white", high = "red", midpoint = 0, na.value = "transparent", name = "Excess \nmortality", limits = c(-6.6, 6.6)) +
  labs(x= "Flu season", y = "Age", tag = "Lower \nmortality") +
  theme_minimal() + 
   theme(plot.tag.position = c(0.8875, 0.375),
        plot.tag = element_text(hjust = 0, size = 15),
        axis.text.x = element_text(angle = 60, hjust = 1, size = 15),
        axis.text.y = element_text(size = 15),
        text = element_text(size = 15, family = "Helvetica"),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=1)) + 
  scale_x_continuous(expand = c(0, 0), breaks = seq(1970, 2020, by = 5), 
                     labels = c("1970", "1975", "1980", "1985", "1990", "1995", "2000",
                                "2005",  "2010", "2015", "2020")) + 
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10))

ggsave(filename = "~/flu_imprinting/plots/final_figs/pancreatic_cancer_residuals_plot.png", 
       plot = p_panc, height = 6, width = 8, dpi = 800, bg='#ffffff')
```


residual plot of ax and age profile. not used?
```{r}
ax_bx$ax_norm <- scale(ax_bx$ax)

colMeans(ax_bx$ax_norm)
apply(ax_bx$ax_norm, 2, sd)

names(ax_bx)[names(ax_bx) == "ax_norm[, 1]"] <- "ax_norm"

mx_small <- mx_data %>% 
  dplyr::select(c(age_recode, year_recode, log_flu_mx)) %>% 
  pivot_wider(names_from = year_recode, values_from = log_flu_mx)

mx_small <- data.frame(mx_small)

rownames(mx_small) <- mx_small[,1]
mx_small <- mx_small[,-1]

mx_small <- mx_small %>% 
  mutate(across(everything(), ~ ifelse(is.na(.), mean(., na.rm = TRUE), .)))

mx_small <- data.frame(scale(mx_small)) %>% 
  mutate(age_recode = as.numeric(rownames(mx_small))) %>%
  pivot_longer(cols = -age_recode,
               names_to = "year_recode",
               values_to = "log_flu_mx_norm") %>% 
  left_join(ax_bx[, c("age_recode", "ax_norm")], by = "age_recode") %>%
  mutate(year_recode = as.numeric(sub("^X", "", year_recode)))

mx_small$resid <- mx_small$log_flu_mx_norm - mx_small$ax_norm
```

```{r}
p=ggplot(mx_small,
       aes(x = year_recode, y = age_recode, fill = resid)) +
  geom_tile() +
  scale_fill_gradient2(low = "darkblue", mid = "white", high = "red", midpoint = 0, na.value = "transparent", name = "Excess \nmortality") +
labs(x= "Flu season", y = "Age", tag = "Lower \nmortality") +
  theme_minimal() + 
  theme(plot.tag.position = c(0.912, 0.41),
        plot.tag = element_text(hjust = 0, size = 15),
        axis.text.x = element_text(angle = 60, hjust = 1, size = 15),
        axis.text.y = element_text(size = 15),
        text = element_text(size = 15, family = "arial"),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=1)) + 
  scale_x_continuous(expand = c(0, 0), breaks = seq(1970, 2020, by = 5), 
                     labels = c("1970", "1975", "1980", "1985", "1990", "1995", "2000",
                                "2005",  "2010", "2015", "2020")) + 
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10))

ggsave(filename = "~/flu_imprinting/plots/final_figs/ax_residuals_plot.png", 
       plot = p, height = 8, width = 10, dpi = 800, bg='#ffffff')
```




part of table 1. manually added in birth years prior to 1918
```{r}
tab1 <- mx_data %>% 
  filter(year_recode == 2020) %>% 
  group_by(imprinted_strain) %>% 
  summarise(
    num_birth_years = n_distinct(birth_year),
    birth_years = paste(sort(unique(birth_year)), collapse = ", ")
  )

print(tab1)

```
